---
title: "Games under <br> Asymmetric Information"
author: "Paulo Fagandini"
institute: "Prague University of Economics and Business <br> Faculty of Informatics and Statistics"
format: 
  revealjs:
    self-contained: false
    logo: ../images/vse.jpg
    slide-number: true
    cross-ref: true
    incremental: true
    math: 
      method: mathjax
    transition: slide
    footer: "Economics of Incentives"
    from: markdown+emoji
    code-overflow: wrap
    code-block-font-size: 0.75em
    progress: true
editor: source

---

<div style="font-size: 70%;">
</div>

# Moral Hazard and Incentives Theory

## Moral Hazard

:::{.callout-note}

## Definition

**Moral Hazard** is a term that refers to the situation when an person acts when she does not bear the full (good or bad) consequences of her acts.

:::

For example if you own a :car:, and if you crash it, you will have to pay all the costs with fixing it :wrench:.

Now, if you have insurance, these costs pass along to the insurer. It is expected that you would drive less carefully than without insurance.

## Moral Hazard

The true cause of the problem is that the Insurance Company cannot know, at every moment, how are you driving. Of course there are some things that are present in the contract:

1. They will not pay if you were speeding.
2. They will not pay if you were driving under the influence of drugs.
3. etc.

But there are a plethora of irresponsible behavior that the company cannot monitor.

## Moral Hazard

In this case, the person who has to *act* and has *private information* on his actions, is known as the **agent**, while who is bearing the consequences of the actions of the agent is called the **principal**.

At the core this is a problem of missaligned incentives: what is good for the agent, might not be necessarily good for the principal.

## Moral Hazard

What to do then?

Principals have to design schemes that *incentive* the agent to behave more in line with the Principal's goals. Of course, these incentives are costly.

## A Principal-Agent model

Suppose the following alternative setup:

1. $e$ is effort exherted by the Agent, with cost $c(e)=\frac{e^2}{2}$.
2. Let $x(e)=x\cdot e$ the return expected for the Principal.
3. Suppose the contract is based on a compensation $w(x)=a+bx$. Note that $b\leq 1$, otherwise the Principal loses money, it would be giving away everything and then some.
4. Suppose the Agent can get utility $\underline{u}$ in an alternative job.

## First Best

The Principal **observes** $e$. The contract must be such that:

1. The agent *accepts* to sign the contract: $$a+bxe-\frac{e^2}{2}\geq \underline{u}$$ this is the **participation constraint**.
2. Maximizes the Principal's expected utility, *i.e.* it solves: $$\max_{e,a,b}\ xe -a-bxe$$

## First Best

Note that the Principal has no incentive to give more than the strictly necessary to make the Agent sign the contract (first best, $e$ is observable), then the participation constraint will be binding.

$$a+bxe-\frac{e^2}{2}= \underline{u}$$

We can substitute this into the Principal's maximization problem.

## First Best

$$\max_{e}\ xe-\frac{e^2}{2}$$

With FOC:

$$x-e^*=0\quad\Rightarrow\quad e^*=x$$

This is the socially optimal level of effort, which will yield an output $x^2$ and set $a$ and $b$ such that $a+bx^2=\underline{u}+\frac{x^2}{2}$. In particular $a=\underline{u}$ and $b=\frac{1}{2}$ would work.

## Second Best

Now assume the Principal cannot observe $e$. What will the Agent do when offered a contract $a+bx(e)$?

. . .

The agent will solve:

$$\max_{\hat{e}} a+bx(\hat{e})-\frac{\hat{e}^2}{2}$$

And choose $e = bx < x$. Note that the larger the bonus, the more effor the Agent will exhert! $a$ seems to be irrelevant for the amount of effort exherted.

## Second Best

This is an anomaly, as people usually are not risk neutral. Let's assume $u()$ such that $u'>0$ and $u''<0$. Now the problem becomes:

$$\max_{\hat{e}} u(a+bx\hat{e})-\frac{\hat{e}^2}{2}$$

And the FOC will be: $$e=u'(a+bxe)bx$$ now we can see that a larger value of $a$ will actually *decrease* the level of effort. Now let's go back to the risk neutral case.

## Second Best

Of course, the Principal will know the Agent is planning her effort level as a function of $a$ and $b$, so she needs to be careful setting those values 

$$
\begin{aligned}
  \max_{e,a,b}\ &xe-a+-bxe\\
  \text{s.t.}\quad &a+bxe-\frac{e^2}{2}=\underline{u}\\
  &e\in\arg\max_{\hat{e}} a + bxe - \frac{e^2}{2}
\end{aligned}
$$

This last equation is the **incentive compatiblity** constraint.

## Second Best

Substituting the IC for its solution and rearranging we have:

$$
\begin{aligned}
  \max_{e,a,b}\ &-a+(1-b)xe\\
  \text{s.t.}\quad &a+bxe-\frac{e^2}{2}=\underline{u}\\
  &e=bx
\end{aligned}
$$

We can substitute the $IC$ in the maximization problem and $PC$ to get

## Second Best

And substituting in the maximization problem yields

$$
\begin{aligned}
  \max_{a,b}\ &-a+(1-b)bx^2\\
  \text{s.t.}\quad &a+\frac{(bx)^2}{2}=\underline{u}
\end{aligned}
$$

And further substituting the $PC$ in the maximization problem we have:

## Second Best

$$
  \max_{b}\ -\left(\underline{u}-\frac{(bx)^2}{2}\right)+(1-b)bx^2
$$

which is equivalent to:

$$
  \max_{b}\ -\underline{u}+\left(1-\frac{b}{2}\right)bx^2
$$

With $FOC:$ $x^2=b^*x^2$, or $b^*=1$. What does this mean?

## Second Best

Remember we discussed briefly the situation of a utility function $u()$ with $u'()>0$ and $u''<0$ for the agent? Why just for the agent?

. . .

We could do that for the principal as well, but usually we can assume firms as risk neutral because shareholders can get rid of the risk by building a diversified portfolio. This is not the case for the agent. We will think on both cases.

## Second Best
<div style="font-size: 70%;">

Let's assume now that:

1. Returns to effort are $x(e)$ is distributed with mean $xe$ and variance $s^2$.
2. If the contract establishes $a$ as base salary and $b$ as variable pay, the wage is a random variable with mean $a+bxe$ and variance $b^2s^2$.
3. Suppose the agent is risk averse, so the utility function is now $$u=a+bxe-\frac{Ke^2}{2}-\frac{\rho b^2s^2}{2}$$ Now $K$ allow us to measure the cost of the effort, and $\rho$ captures the level of risk aversion of the agent.
  
</div>

## Second Best

Now, given $a$ and $b$, what is hte optimal level of effort?

$$e^*=\frac{bx}{K}$$

The only difference, is that the most costly effort is, the lower the optimal level of effort for the same compensation. [Homework: Introduce $K$ in the previous solution and check that the solution is the same]

## Second Best

The principal's problem is:

$$
\begin{aligned}
  \max_{e,a,b}\ &xe-a+-bxe\\
  \text{s.t.}\quad &a+bxe-\frac{Ke^2}{2}-\frac{\rho b^2s^2}{2}=\underline{u}\\
  &e\in\arg\max_{\hat{e}} a + bxe - \frac{e^2}{2}-\frac{\rho b^2s^2}{2}
\end{aligned}
$$

Again, we can replace the $IC$


## Second Best

$$
\begin{aligned}
  \max_{e,a,b}\ &xe-a+-bxe\\
  \text{s.t.}\quad &a+bxe-\frac{Ke^2}{2}-\frac{\rho b^2s^2}{2}=\underline{u}\\
  &e=\frac{bx}{K}
\end{aligned}
$$

And we can replace the $IC$ the the $PC$ and the maximization problem.

## Second Best

$$
\begin{aligned}
  \max_{a,b}\ &-a+(1-b)\frac{bx^2}{K}\\
  \text{s.t.}\quad &a+\frac{(bx)^2}{K}-\frac{(bx)^2}{2K}-\frac{\rho b^2s^2}{2}=\underline{u}
\end{aligned}
$$

## Second Best

$$
\begin{aligned}
  \max_{a,b}\ &-a+(1-b)\frac{bx^2}{K}\\
  \text{s.t.}\quad &a+\frac{(bx)^2}{2K}-\frac{\rho b^2s^2}{2}=\underline{u}
\end{aligned}
$$
And further substituting $a$ into the maximization problem we obtain:

## Second Best

$$
\begin{aligned}
  \max_{b}\ &-\left(\underline{u}-\frac{(bx)^2}{2K}+\frac{\rho b^2s^2}{2}\right)+(1-b)\frac{bx^2}{K}
\end{aligned}
$$

## Second Best

With FOC: $$\frac{bx^2}{K}-\rho s^2 b+\frac{(1-2b)x^2}{K}=0$$
$$bx^2-\rho s^2Kb+(1-2b)x^2=0$$
$$-\rho s^2Kb+(1-b)x^2=0$$

## Second Best

$$b^*=\frac{x^2}{\rho s^2 K + x^2}$$

Note this is:

1. Increasing in $x^2$
2. Decreasing in $K$
3. Decreasing in $\rho$ and $s^2$.

# Games with Incomplete Information

## Incomplete Information

Games where players are uncertain about:

- Payoffs of other players
- Available strategies of opponents
- Types of players they face

. . .

Key distinction: Players know their own type but are uncertain about others' types

## Example 1

Consider two candidates campaigning for some election

* Each player can run a "clean" campaign, or
* Each player runs a "dirty" campaign

. . .

| P1-P2 | dirty | clean |
| :--: | :--: | :--: |
| dirty | (0,0) | (7,-2) |
| clean | (-2,7) | (5,5) |

## Example 1

Now suppose that:

* Player 2 is as we know, for him, playing *dirty* is a dominant strategy.
* Player 1's preferences are unknown to player 2, he can be *tough* (he is like P2) or *accomodanting* (*clean* is dominant).
* P1 knows exactly the payoff matrix (he knows his type), but P2 does not.

## Example 1

:::{.columns}

:::{.column}

| P1-P2 | dirty | clean |
| :--: | :--: | :--: |
| dirty | (0,0) | (7,-2) |
| clean | (-2,7) | (5,5) |

P1 is *tough*
:::

:::{.column}

| P1-P2 | dirty | clean |
| :--: | :--: | :--: |
| dirty | (-2,0) | (5,-2) |
| clean | (0,7) | (7,5) |

P1 is *accomodating*
:::

:::

Because this game has dominant strategies, the $NE$ is easy to spot, $(d,dc)$, *i.e.*, each player (type) plays his dominant strategy.

## Example 1

```{r, engine='tikz', fig.align='center', clean=TRUE}

\usetikzlibrary{calc,fit,shapes.geometric}

\tikzset{
  solid node/.style={circle,draw,inner sep=1.5,fill=black},
  hollow node/.style={circle,draw,inner sep=1.5}
}

\begin{tikzpicture}[scale=1.5,font=\tiny]
  \tikzstyle{level 1}=[level distance=10mm,sibling distance=30mm]
  \tikzstyle{level 2}=[level distance=10mm,sibling distance=15mm]
  \tikzstyle{level 3}=[level distance=10mm,sibling distance=7.5mm]

  \node(0)[solid node,label=above:{$N$}]{}
    child{
      node(1)[solid node, label=above left:{$P1$}]{}
        child{
          node(3)[solid node, label=above left:{}]{}
            child{
              node(7)[hollow node, label=below:{$(0,0)$}]{}
              edge from parent node[left]{$d$}
            }
            child{
              node(8)[hollow node, label=below:{$(7,-2)$}]{}
              edge from parent node[right]{$c$}
            }
          edge from parent node[left]{$d$}
        }
        child{
          node(4)[solid node, label=above right:{}]{}
            child{
              node(9)[hollow node, label=below:{$(-2,7)$}]{}
              edge from parent node[left]{$d$}
            }
            child{
              node(10)[hollow node, label=below:{$(5,5)$}]{}
              edge from parent node[right]{$c$}
            }
          edge from parent node[right]{$c$}
        }
      edge from parent node[left]{$tough$}
    }
    child{
      node(2)[solid node, label=above right:{$P1$}]{}
      child{
          node(5)[solid node, label=above left:{}]{}
            child{
              node(11)[hollow node, label=below:{$(-2,0)$}]{}
              edge from parent node[left]{$d$}
            }
            child{
              node(12)[hollow node, label=below:{$(5,-2)$}]{}
              edge from parent node[right]{$c$}
            }
          edge from parent node[left]{$d$}
        }
        child{
          node(6)[solid node, label=above right:{}]{}
            child{
              node(13)[hollow node, label=below:{$(2,7)$}]{}
              edge from parent node[left]{$d$}
            }
            child{
              node(14)[hollow node, label=below:{$(7,5)$}]{}
              edge from parent node[right]{$c$}
            }
          edge from parent node[right]{$c$}
        }
      edge from parent node[right]{$accomodating$}
    };

   \draw[dashed,rounded corners=10]($(3) + (-.2,.25)$)rectangle($(6) +(.2,-.25)$);
   \node at ($(3)!.5!(6)$) {$P2$};

\end{tikzpicture}

```

## Example 1

As expected, this is the simplest scenario we could face, an equlibrium with dominant strategies. This will not be the case, and we will need to study the *beliefs* of players regarding the other player's **type**.

## Example 2

<div style="font-size: 70%;">

1. Suppose there are two planets trying to settle their diplomatic relationships, **H**oth and **W**obani.
2. They have to decide if they will be sign a **F**ree Trade Agreement or they will sign a  **S**ecurity pact.
3. **H**oth is a pacifist country, and prefers to focus in trade, while **H**oth is concerned with their security, and then they would prefer a security pact.
4. **W**obani's diplomatic team can be one of two types, *integrationist* or *isolationist*. This is **unknown** to **H**oth's diplomatic team.

</div>

## Example 2

:::{.columns}

:::{.column}

| H - W | $F$ | $S$ |
| :--: | :--: | :--: |
| $F$ | (3,1) | (0,0) |
| $S$ | (0,0) | (1,3) |

**W** is Integrationist (type 1)

:::

:::{.column}

| H - W | $F$ | $S$ |
| :--: | :--: | :--: |
| $F$ | (3,0) | (0,1) |
| $S$ | (0,3) | (1,0) |

**W** is Isolationist. (type 2)

:::

:::

## Example 2

1. **W** knows exactly her type.
2. Given that **H** does not know the type of **W**, it will a probability $p$ to the scenario in which **W** is integrationist.
3. Let's assume that **W** knows $p$.

. . .

The method to solve these games was developed by the Hungarian economist John Harsanyi.

## Example 2

1. Turn this game into an Imperfect Information Game. Introduce $N$ as a player.
2. Use the $NE$ concept we used for imperfect information games (SPNE). This will be called **Bayes Nash Equilibrium**.

## Bayesian Games Framework

<div style="font-size: 85%;">

A Bayesian game consists of:

:::{.nonincremental}

- Set of players $N = \{1, 2, ..., n\}$
- Type spaces $T_i$ for each player $i$
- Prior beliefs $p(t)$ over type profiles
- Action spaces $A_i$ for each player
- Payoff functions $u_i(a, t)$ depending on actions and types

:::

Players maximize expected utility given beliefs about opponents' types

</div>

## Bayesian Nash Equilibrium

A strategy profile where each player's strategy is optimal given:

:::{.nonincremental}

- Their beliefs about other players' types
- The strategies of other players

:::

Formally: $s_i^*(t_i) \in \arg\max_{a_i} \sum_{t_{-i}} p(t_{-i}|t_i) u_i(a_i, s_{-i}^*(t_{-i}), t_i, t_{-i})$

Each type of each player best responds to opponents' strategies in *expected utility*

## Example 2

<div style="font-size: 85%;">

In this case:

1. A strategy for **H** is:
   1. pure strategies, just $F$ or $S$
   2. mixed strategies, $\lambda\in(0,1)$ where $\lambda$ is the probability of playing $F$.
2. A strategy for **W** is:
   1. pure strategies, a pair of $F$ or $S$ for each type.
   2. mixed stragies, a pair $(\mu_1,\mu_2)$ for the probability of playing $F$ for each type.

</div>

## Example 2

A Bayes Nash Equilbrium (BNE) will be a triple $(\lambda, \mu_1,\mu_2)$ in which each player (and each type of player) will play a best response as follows:

:::{.nonincremental}

1. $\mu_i$ maximizes type's $i$ **W** expected payoffs, if **H** choses $F$ with probability $\lambda$.
2. $\lambda$ maximizes **H**'s expected payoffs if **H** believes that with probability $p$ she will encounter **W** type 1, that will play $F$ with probability $\mu_1$.

:::

## Example 2

```{r, engine='tikz', fig.align='center', clean=TRUE}

\usetikzlibrary{calc,fit,shapes.geometric}

\tikzset{
  solid node/.style={circle,draw,inner sep=1.5,fill=black},
  hollow node/.style={circle,draw,inner sep=1.5}
}

\begin{tikzpicture}[scale=1.5,font=\tiny]
  \tikzstyle{level 1}=[level distance=10mm,sibling distance=30mm]
  \tikzstyle{level 2}=[level distance=10mm,sibling distance=15mm]
  \tikzstyle{level 3}=[level distance=10mm,sibling distance=7.5mm]

  \node(0)[solid node,label=above:{$N$}]{}
    child{
      node(1)[solid node, label=above left:{$Wobani$}]{}
        child{
          node(3)[solid node, label=above left:{}]{}
            child{
              node(7)[hollow node, label=below:{$(1,3)$}]{}
              edge from parent node[left]{$F$}
            }
            child{
              node(8)[hollow node, label=below:{$(0,0)$}]{}
              edge from parent node[right]{$S$}
            }
          edge from parent node[left]{$F$}
        }
        child{
          node(4)[solid node, label=above right:{}]{}
            child{
              node(9)[hollow node, label=below:{$(0,0)$}]{}
              edge from parent node[left]{$F$}
            }
            child{
              node(10)[hollow node, label=below:{$(3,1)$}]{}
              edge from parent node[right]{$S$}
            }
          edge from parent node[right]{$S$}
        }
      edge from parent node[left]{$Type 1$}
    }
    child{
      node(2)[solid node, label=above right:{$Wobani$}]{}
      child{
          node(5)[solid node, label=above left:{}]{}
            child{
              node(11)[hollow node, label=below:{$(0,3)$}]{}
              edge from parent node[left]{$F$}
            }
            child{
              node(12)[hollow node, label=below:{$(3,0)$}]{}
              edge from parent node[right]{$S$}
            }
          edge from parent node[left]{$F$}
        }
        child{
          node(6)[solid node, label=above right:{}]{}
            child{
              node(13)[hollow node, label=below:{$(1,0)$}]{}
              edge from parent node[left]{$F$}
            }
            child{
              node(14)[hollow node, label=below:{$(0,1)$}]{}
              edge from parent node[right]{$S$}
            }
          edge from parent node[right]{$S$}
        }
      edge from parent node[right]{$Type 2$}
    };

   \draw[dashed,rounded corners=10]($(3) + (-.2,.25)$)rectangle($(6) +(.2,-.25)$);
   \node at ($(3)!.5!(6)$) {$Hoth$};

\end{tikzpicture}

```

## Example 2: Pure Strategies

1. Suppose $H$ plays $F$ for sure.
2. In this case, $W$ will play $FS$.

. . .

But can this be an equilibrium?

. . .

If so, **H** must stick with $F$, $i.e.$ $F$ must be a best response to **W** strategy's $FS$.

## Example 2: Pure Strategies

<div style="font-size: 85%;">

Remember, **H** expects **W** to be type 1 with probability $p$, so her expected payoffs are:

:::{.nonincremental}

1. $E[\pi_H(F,FS)]=p3+(1-p)0=3p$
2. $E[\pi_H(S,FS)]=p0+(1-p)1=1-p$

:::

If **H** is to stick with $F$ when **W** is playing $FS$, then it must be true that:

$$3p\geq 1-p\ \Rightarrow\ p\geq \frac{1}{4}$$

</div>

## Example 2: Pure Strategies

:::{.nonincremental}

1. Suppose now that **H** plays $S$ for sure.
2. In this case, **W** will play $SF$.

:::

But can this be an equilibrium? If so, **H** must stick with $S$, $i.e.$ $S$ must be a best response to **W** strategy's $SF$.

## Example 2: Pure Strategies

<div style="font-size: 85%;">

:::{.nonincremental}

Remember, **H** expects **W** to be type 1 with probability $p$, so her expected payoffs are:

1. $E[\pi_H(F,SF)]=p0+(1-p)3=3(1-p)$
2. $E[\pi_H(S,SF)]=p1+(1-p)0=p$

If **H** is to stick with $S$ when **W** is playing $SF$, then it must be true that:

$$3(1-p)\leq p\ \Rightarrow\ p\geq \frac{3}{4}$$

:::

</div>

## Example 2: Pure Strategies

Conclusion:

1. If $p\geq \frac{3}{4}$ there are 2 BNEs, $(F,FS)$ and $(S,SF)$.
2. If $\frac{1}{4}\leq p < \frac{3}{4}$ there is only 1 BNE, $(F,FS)$.
3. If $p<\frac{1}{4}$ there is no BNE in pure strategies.

## Example 2: Mixed Strategies

:::{.callout-important}
Remember that in the case of mixed strategies, to randomize, each pure strategy with positive probability must be yield the same expected payoff.
:::

There can be a plethora of NEs in mixed strategies, let's focus on a special situation, when $p<\frac{1}{4}$ which is a situation when we know for sure there is no pure strategy $BNE$.

## Example 2: Mixed Strategies

1. Suppose **H** plays $F$ with probability $\lambda$.
2. Type 1 **W** has expected payoff $\lambda$ by playing $F$, while playing $S$ can expect $3(1-\lambda)$.
2. Type 2 **W** has expected payoff by playing $F$ $3(1-\lamdba)$, while playing $S$ can expect $\lambda$.

## Example 2: Mixed Strategies

1. **W** type 1 would play a mixed strategies equilibrium only if $\lambda=3(1-\lambda)$, or $\lambda=\frac{3}{4}$.
2. **W** type 2 would play a mixed strategies equilibrium only if $3(1-\lambda)=\lambda$, or $\lambda=\frac{3}{4}$.

## Example 2: Mixed Strategies

A BNE of this game is $\lambda=\frac{3}{4}$, $(\mu_1,\mu_2)=\left(\frac{1}{4},\frac{1}{4}\right)$

# Mechanism Design, the Revelation Principle, and Sales to an Unknown Buyer

## Mechanism Design

* So far we have studied what players will do, given a set of rules and payoffs (not entirely true, *i.e.* Contracts)
* Mechanism Design asks, if I want a specific outcome:
  * What de rules must be?
  * Is it even possible?

. . .

Basically now we design the game to induce the players to play in a specific manner.

## Mechanism Design

In general, the set of players is something that will be given, *i.e.* is something we cannot **design**, but the **designer** can choose:

1. choices available to players
2. payoffs

## Mechanism Design

However, there are some constraints for the *designer*

1. There cannot be any coercion (why would you need to incentivize behavior if you can force it)
2. Players can choose to not participate, and therefore the game must give enough incentives to make players want to play.

The optimal design is a mechanism whose equilibrium is the one that generates the highest payoff for the designer.

## Mechanism Design

More formally:

:::{.callout}

## Definition

A mechanism is a game (a set of rules) that specifies the strategies that the player can choose from and the outcome for every choice.

:::

## Mechanism Design

1. Let a strategy for a player $s$.
2. Let the outcome of a game be denoted by $t$.
3. Let the *type* of the player be denoted by $\theta$.
4. Let the payoff of the player be represented by the function $\pi(s,t,\theta)$.

The designer can choose $s$ and $t$, but $\theta$ and $\pi$ is out of his control.

## Mechanism Design

:::{.callout}

## Definition

Suppose there are two types $\theta$ and $\mu$. The designer assigns $(s_{\theta},t_{\theta})$ as strategies and outcomes for the player type $\theta$, and $(s_{\mu},t_{\mu})$ for the player type $\mu$.

A design is said to be **incentive compatible** if:

$$\pi(s_{\theta},t_{\theta},\theta)\geq \pi(s,t,\theta)\quad \forall s,\ \forall t$$ and
$$\pi(s_{\mu},t_{\mu},\mu)\geq \pi(s,t,\mu)\quad \forall s,\ \forall t$$

:::

This is, every player has incentives to *tell the truth*.

## Mechanism Design

:::{.callout}

## Definition

Suppose there are two types $\theta$ and $\mu$, with *reservation utility* $\underline{\pi}_{\theta}$ and $\underline{\pi}_{\mu}$.

A design is said to be **idividually rational** if:

$$\pi(s_{\theta},t_{\theta},\theta)\geq \underline{\pi}_\theta$$ and
$$\pi(s_{\mu},t_{\mu},\mu)\geq \underline{\pi}_\mu$$

:::

This is, every player has incentives to *participate*.

## Mechanism Design

Without much surprise we call the designer the **Principal**.

Any design the Principal devices to maximize her expected payoff must be **incentive compatible** and **individually rational**. Her payoff usually depends on the strategies chosen by the players.

For surprise to none, there as infinite number of ways in which a Principal can set a mechanism, however we will focus on the class named **direct-revelation mechanisms**.

## Mechanism Design

Direct-revelation mechanism:

1. Each player is free to lie about his type. $\theta$ can claim to be $\mu$ if so desires, and vice-versa.
2. Suppose that if you claim to be type $\theta$, the mechanism assings you $(s_\theta,t_\theta)$, and $(s_\mu,t_\mu)$ for the $\mu$ type

# Auctions

## Auctions

Auctions are, probably, the most transparent mechanism.

1. Auctioneer does not know the player's individual valuations.
2. Auctioneer wants to exert as much money as she can from bidders.

# Signaling Games and The Lemons Problem



# Bibliography

## Bibliography

These slides took definitions and examples from the book:

:::{.nonincremental}

* Strategies and Games: Theory and Applications from Prajit Dutta. 2001 MIT Press. Chapters 19 to 24.
* The Moral Hazard example was taken from *Wealth and the Principal-Agent Matching*, by P. Fagandini (2022), Managerial and Decision Economics.

:::

