---
title: "Games under <br> Asymmetric Information"
author: "Paulo Fagandini"
institute: "Prague University of Economics and Business <br> Faculty of Informatics and Statistics"
format: 
  revealjs:
    self-contained: false
    logo: ../images/vse.jpg
    slide-number: true
    cross-ref: true
    incremental: true
    math: 
      method: mathjax
    transition: slide
    footer: "Economics of Incentives"
    from: markdown+emoji
    code-overflow: wrap
    code-block-font-size: 0.75em
    progress: true
editor: source

---

<div style="font-size: 70%;">
</div>

# Moral Hazard and Incentives Theory

## Moral Hazard

:::{.callout-note}

## Definition

**Moral Hazard** is a term that refers to the situation when an person acts when she does not bear the full (good or bad) consequences of her acts.

:::

For example if you own a :car:, and if you crash it, you will have to pay all the costs with fixing it :wrench:.

Now, if you have insurance, these costs pass along to the insurer. It is expected that you would drive less carefully than without insurance.

## Moral Hazard

The true cause of the problem is that the Insurance Company cannot know, at every moment, how are you driving. Of course there are some things that are present in the contract:

1. They will not pay if you were speeding.
2. They will not pay if you were driving under the influence of drugs.
3. etc.

But there are a plethora of irresponsible behavior that the company cannot monitor.

## Moral Hazard

In this case, the person who has to *act* and has *private information* on his actions, is known as the **agent**, while who is bearing the consequences of the actions of the agent is called the **principal**.

At the core this is a problem of missaligned incentives: what is good for the agent, might not be necessarily good for the principal.

## Moral Hazard

What to do then?

Principals have to design schemes that *incentive* the agent to behave more in line with the Principal's goals. Of course, these incentives are costly.

## A Principal-Agent model

The most simple way to model the Pricipal-Agent model (intuitive and transparent way) is by modelling the relationship between employer (Principal) and employee (Agent).

Suppose the Principal hires the Agent to perform some task. 

## A Principal-Agent model
Let:

1. both agent and principal be risk neutral: $u_a(w)=w$, and $u_p(\pi)=\pi$.
2. the outcome be $x\in\{0,\xi\}$.
3. effort be chosen as $e\in[0,1]$, and $Pr(x=\xi|e)=e$.
4. the wage have fixed $a$ pay and bonus $b$, paid in case $x=\xi$.
5. Let $c(e)=\frac{e^2}{2}$ be the cost of effort.

## A Principal-Agent model

The maximization problem of the principal is given by:

$$
\begin{aligned}
  \max_{e,a,b}\quad &-a+e(\xi-b)\\
  s.t.\quad &a+eb-c(e)\geq\bar{u}\\
  &e\in\arg\max_{\hat{e}}\{a+\hat{e}b-c(\hat{e})\}\\
  &a\geq -w
\end{aligned}
$$

Let's assume, further, than $\omega = 0$, *i.e.* we will not allow negative fixed wage (or transfers from the agent to the principal). This is know as limited liability or cash constraint.

## A Principal-Agent model

Grossman and Hart (1983) show that under some conditions (satisfied in this model and in most of the simple versions of this model) the problem can be solved in two stages:

1. Solve the problem for the agent, taking everything external to the agent as a parameter.
2. Solve the problem of the principal, substituting the solution of the agent as a function of the principal's variables.

This is how we will proceed.

## A Principal-Agent model

Given $a$ and $b$, the agent will maximize her utility, so computing the FOC we get:

$$b-c'=0\quad\Rightarrow\quad b=e^*$$

So the agent will choose optimally $e=b$.

## A Principal-Agent model

The problem for the principal is now:

$$
\begin{aligned}
  \max_{a,e}\quad &-a+e(\xi-e)\\
  s.t.\quad &a+e^2-\frac{e^2}{2}\geq\bar{u}\\
  &a\geq 0
\end{aligned}
$$

Now, if $e^*=0$, the agent will accept to participate if and only if $a\geq \bar{u}$. For simplicity assume $\bar{u}=0$.

## A Principal-Agent model

$$
\begin{aligned}
  \max_{a,e}\quad &-a+e(\xi-e)\\
  s.t.\quad &a+\frac{e^2}{2}\geq0\\
  &a\geq 0
\end{aligned}
$$

Which constraint is irrelevant now?

. . .

The Participation Constraint, as $\frac{e^2}{2}\geq 0$. Then the optimal is $a=0$. Now what is the optimal level of effort?

## A Principal-Agent model

$$
\begin{aligned}
  \max_{e}\quad &e(\xi-e)\\
\end{aligned}
$$

And from there $e^*=\frac{\xi}{2}$. Note that this makes $b=\frac{\xi}{2}$. So the expected payoffs are:

1. $E[u_a]=0+\left(\frac{\xi}{2}\right)^2-\frac{\xi^2}{8}=\frac{\xi^2}{8}$
2. $E[u_p]=-0+\frac{\xi}{2}\left(\xi-\frac{\xi}{2}\right)=\frac{\xi^2}{4}$

## A Principal-Agent model

What if the Principal could observe effort?

Then the Incentive Compatibility is unnecessary, and the principal needs just to compensate effort (non probabilistically)

$$
\begin{aligned}
  \max_{e}\quad &e\xi-c(e)\\
\end{aligned}
$$

Which has solution $e^*=\xi$, $a=0$, and $b=c(\xi)=\frac{\xi^2}{2}$. The expected utility of the agent is $0$, and the expected profit for the principal is $\frac{\xi^2}{2}$.

## A Principal-Agent model

Some notes:

* With information asymmetry, the agent can *extract information rents* from the principal, at the cost of a worse outcome overall.
* Without information asymmetry, the principal can extrat the whole surplus, which is larger than the case with information asymmetries.
  
## A Principal-Agent model

In general,

1. Incentivizing hard work is **costly**.
2. The higher the profits, the larger the bonus.
3. The principal is almost always worse off in the presence of moral hazard, while the agent might be better off.

# Games with Incomplete Information

## Incomplete Information

Games where players are uncertain about:

- Payoffs of other players
- Available strategies of opponents
- Types of players they face

. . .

Key distinction: Players know their own type but are uncertain about others' types

## Example 1

Consider two candidates campaigning for some election

* Each player can run a "clean" campaign, or
* Each player runs a "dirty" campaign

. . .

| P1-P2 | dirty | clean |
| :--: | :--: | :--: |
| dirty | (0,0) | (7,-2) |
| clean | (-2,7) | (5,5) |

## Example 1

Now suppose that:

* Player 2 is as we know, for him, playing *dirty* is a dominant strategy.
* Player 1's preferences are unknown to player 2, he can be *tough* (he is like P2) or *accomodanting* (*clean* is dominant).
* P1 knows exactly the payoff matrix (he knows his type), but P2 does not.

## Example 1

:::{.columns}

:::{.column}

| P1-P2 | dirty | clean |
| :--: | :--: | :--: |
| dirty | (0,0) | (7,-2) |
| clean | (-2,7) | (5,5) |

P1 is *tough*
:::

:::{.column}

| P1-P2 | dirty | clean |
| :--: | :--: | :--: |
| dirty | (-2,0) | (5,-2) |
| clean | (0,7) | (7,5) |

P1 is *accomodating*
:::

:::

Because this game has dominant strategies, the $NE$ is easy to spot, $(d,dc)$, *i.e.*, each player (type) plays his dominant strategy.

## Example 1

```{r, engine='tikz', fig.align='center'}

\usetikzlibrary{calc,fit,shapes.geometric}

\tikzset{
  solid node/.style={circle,draw,inner sep=1.5,fill=black},
  hollow node/.style={circle,draw,inner sep=1.5}
}

\begin{tikzpicture}[scale=1.5,font=\tiny]
  \tikzstyle{level 1}=[level distance=10mm,sibling distance=30mm]
  \tikzstyle{level 2}=[level distance=10mm,sibling distance=15mm]
  \tikzstyle{level 3}=[level distance=10mm,sibling distance=7.5mm]

  \node(0)[solid node,label=above:{$N$}]{}
    child{
      node(1)[solid node, label=above left:{$P1$}]{}
        child{
          node(3)[solid node, label=above left:{}]{}
            child{
              node(7)[hollow node, label=below:{$(0,0)$}]{}
              edge from parent node[left]{$d$}
            }
            child{
              node(8)[hollow node, label=below:{$(7,-2)$}]{}
              edge from parent node[right]{$c$}
            }
          edge from parent node[left]{$d$}
        }
        child{
          node(4)[solid node, label=above right:{}]{}
            child{
              node(9)[hollow node, label=below:{$(-2,7)$}]{}
              edge from parent node[left]{$d$}
            }
            child{
              node(10)[hollow node, label=below:{$(5,5)$}]{}
              edge from parent node[right]{$c$}
            }
          edge from parent node[right]{$c$}
        }
      edge from parent node[left]{$tough$}
    }
    child{
      node(2)[solid node, label=above right:{$P1$}]{}
      child{
          node(5)[solid node, label=above left:{}]{}
            child{
              node(11)[hollow node, label=below:{$(-2,0)$}]{}
              edge from parent node[left]{$d$}
            }
            child{
              node(12)[hollow node, label=below:{$(5,-2)$}]{}
              edge from parent node[right]{$c$}
            }
          edge from parent node[left]{$d$}
        }
        child{
          node(6)[solid node, label=above right:{}]{}
            child{
              node(13)[hollow node, label=below:{$(2,7)$}]{}
              edge from parent node[left]{$d$}
            }
            child{
              node(14)[hollow node, label=below:{$(7,5)$}]{}
              edge from parent node[right]{$c$}
            }
          edge from parent node[right]{$c$}
        }
      edge from parent node[right]{$accomodating$}
    };

   \draw[dashed,rounded corners=10]($(3) + (-.2,.25)$)rectangle($(6) +(.2,-.25)$);
   \node at ($(3)!.5!(6)$) {$P2$};

\end{tikzpicture}

```

## Example 1

As expected, this is the simplest scenario we could face, an equlibrium with dominant strategies. This will not be the case, and we will need to study the *beliefs* of players regarding the other player's **type**.

## Example 2

<div style="font-size: 70%;">

1. Suppose there are two planets trying to settle their diplomatic relationships, **H**oth and **W**obani.
2. They have to decide if they will be sign a **F**ree Trade Agreement or they will sign a  **S**ecurity pact.
3. **H**oth is a pacifist country, and prefers to focus in trade, while **H**oth is concerned with their security, and then they would prefer a security pact.
4. **W**obani's diplomatic team can be one of two types, *integrationist* or *isolationist*. This is **unknown** to **H**oth's diplomatic team.

</div>

## Example 2

:::{.columns}

:::{.column}

| H - W | $F$ | $S$ |
| :--: | :--: | :--: |
| $F$ | (3,1) | (0,0) |
| $S$ | (0,0) | (1,3) |

**W** is Integrationist (type 1)

:::

:::{.column}

| H - W | $F$ | $S$ |
| :--: | :--: | :--: |
| $F$ | (3,0) | (0,1) |
| $S$ | (0,3) | (1,0) |

**W** is Isolationist. (type 2)

:::

:::

## Example 2

1. **W** knows exactly her type.
2. Given that **H** does not know the type of **W**, it will a probability $p$ to the scenario in which **W** is integrationist.
3. Let's assume that **W** knows $p$.

. . .

The method to solve these games was developed by the Hungarian economist John Harsanyi.

## Example 2

1. Turn this game into an Imperfect Information Game. Introduce $N$ as a player.
2. Use the $NE$ concept we used for imperfect information games (SPNE). This will be called **Bayes Nash Equilibrium**.

## Bayesian Games Framework

<div style="font-size: 85%;">

A Bayesian game consists of:

:::{.nonincremental}

- Set of players $N = \{1, 2, ..., n\}$
- Type spaces $T_i$ for each player $i$
- Prior beliefs $p(t)$ over type profiles
- Action spaces $A_i$ for each player
- Payoff functions $u_i(a, t)$ depending on actions and types

:::

Players maximize expected utility given beliefs about opponents' types

</div>

## Bayesian Nash Equilibrium

A strategy profile where each player's strategy is optimal given:

:::{.nonincremental}

- Their beliefs about other players' types
- The strategies of other players

:::

Formally: $s_i^*(t_i) \in \arg\max_{a_i} \sum_{t_{-i}} p(t_{-i}|t_i) u_i(a_i, s_{-i}^*(t_{-i}), t_i, t_{-i})$

Each type of each player best responds to opponents' strategies in *expected utility*

## Example 2

<div style="font-size: 85%;">

In this case:

1. A strategy for **H** is:
   1. pure strategies, just $F$ or $S$
   2. mixed strategies, $\lambda\in(0,1)$ where $\lambda$ is the probability of playing $F$.
2. A strategy for **W** is:
   1. pure strategies, a pair of $F$ or $S$ for each type.
   2. mixed stragies, a pair $(\mu_1,\mu_2)$ for the probability of playing $F$ for each type.

</div>

## Example 2

A Bayes Nash Equilbrium (BNE) will be a triple $(\lambda, \mu_1,\mu_2)$ in which each player (and each type of player) will play a best response as follows:

:::{.nonincremental}

1. $\mu_i$ maximizes type's $i$ **W** expected payoffs, if **H** choses $F$ with probability $\lambda$.
2. $\lambda$ maximizes **H**'s expected payoffs if **H** believes that with probability $p$ she will encounter **W** type 1, that will play $F$ with probability $\mu_1$.

:::

## Example 2

```{r, engine='tikz', fig.align='center'}

\usetikzlibrary{calc,fit,shapes.geometric}

\tikzset{
  solid node/.style={circle,draw,inner sep=1.5,fill=black},
  hollow node/.style={circle,draw,inner sep=1.5}
}

\begin{tikzpicture}[scale=1.5,font=\tiny]
  \tikzstyle{level 1}=[level distance=10mm,sibling distance=30mm]
  \tikzstyle{level 2}=[level distance=10mm,sibling distance=15mm]
  \tikzstyle{level 3}=[level distance=10mm,sibling distance=7.5mm]

  \node(0)[solid node,label=above:{$N$}]{}
    child{
      node(1)[solid node, label=above left:{$Wobani$}]{}
        child{
          node(3)[solid node, label=above left:{}]{}
            child{
              node(7)[hollow node, label=below:{$(1,3)$}]{}
              edge from parent node[left]{$F$}
            }
            child{
              node(8)[hollow node, label=below:{$(0,0)$}]{}
              edge from parent node[right]{$S$}
            }
          edge from parent node[left]{$F$}
        }
        child{
          node(4)[solid node, label=above right:{}]{}
            child{
              node(9)[hollow node, label=below:{$(0,0)$}]{}
              edge from parent node[left]{$F$}
            }
            child{
              node(10)[hollow node, label=below:{$(3,1)$}]{}
              edge from parent node[right]{$S$}
            }
          edge from parent node[right]{$S$}
        }
      edge from parent node[left]{$Type 1$}
    }
    child{
      node(2)[solid node, label=above right:{$Wobani$}]{}
      child{
          node(5)[solid node, label=above left:{}]{}
            child{
              node(11)[hollow node, label=below:{$(0,3)$}]{}
              edge from parent node[left]{$F$}
            }
            child{
              node(12)[hollow node, label=below:{$(3,0)$}]{}
              edge from parent node[right]{$S$}
            }
          edge from parent node[left]{$F$}
        }
        child{
          node(6)[solid node, label=above right:{}]{}
            child{
              node(13)[hollow node, label=below:{$(1,0)$}]{}
              edge from parent node[left]{$F$}
            }
            child{
              node(14)[hollow node, label=below:{$(0,1)$}]{}
              edge from parent node[right]{$S$}
            }
          edge from parent node[right]{$S$}
        }
      edge from parent node[right]{$Type 2$}
    };

   \draw[dashed,rounded corners=10]($(3) + (-.2,.25)$)rectangle($(6) +(.2,-.25)$);
   \node at ($(3)!.5!(6)$) {$Hoth$};

\end{tikzpicture}

```

## Example 2: Pure Strategies

1. Suppose $H$ plays $F$ for sure.
2. In this case, $W$ will play $FS$.

. . .

But can this be an equilibrium?

. . .

If so, **H** must stick with $F$, $i.e.$ $F$ must be a best response to **W** strategy's $FS$.

## Example 2: Pure Strategies

<div style="font-size: 85%;">

Remember, **H** expects **W** to be type 1 with probability $p$, so her expected payoffs are:

:::{.nonincremental}

1. $E[\pi_H(F,FS)]=p3+(1-p)0=3p$
2. $E[\pi_H(S,FS)]=p0+(1-p)1=1-p$

:::

If **H** is to stick with $F$ when **W** is playing $FS$, then it must be true that:

$$3p\geq 1-p\ \Rightarrow\ p\geq \frac{1}{4}$$

</div>

## Example 2: Pure Strategies

:::{.nonincremental}

1. Suppose now that **H** plays $S$ for sure.
2. In this case, **W** will play $SF$.

:::

But can this be an equilibrium? If so, **H** must stick with $S$, $i.e.$ $S$ must be a best response to **W** strategy's $SF$.

## Example 2: Pure Strategies

<div style="font-size: 85%;">

:::{.nonincremental}

Remember, **H** expects **W** to be type 1 with probability $p$, so her expected payoffs are:

1. $E[\pi_H(F,SF)]=p0+(1-p)3=3(1-p)$
2. $E[\pi_H(S,SF)]=p1+(1-p)0=p$

If **H** is to stick with $S$ when **W** is playing $SF$, then it must be true that:

$$3(1-p)\leq p\ \Rightarrow\ p\geq \frac{3}{4}$$

:::

</div>

## Example 2: Pure Strategies

Conclusion:

1. If $p\geq \frac{3}{4}$ there are 2 BNEs, $(F,FS)$ and $(S,SF)$.
2. If $\frac{1}{4}\leq p < \frac{3}{4}$ there is only 1 BNE, $(F,FS)$.
3. If $p<\frac{1}{4}$ there is no BNE in pure strategies.

## Example 2: Mixed Strategies

:::{.callout-important}
Remember that in the case of mixed strategies, to randomize, each pure strategy with positive probability must be yield the same expected payoff.
:::

There can be a plethora of NEs in mixed strategies, let's focus on a special situation, when $p<\frac{1}{4}$ which is a situation when we know for sure there is no pure strategy $BNE$.

## Example 2: Mixed Strategies

1. Suppose **H** plays $F$ with probability $\lambda$.
2. Type 1 **W** has expected payoff $\lambda$ by playing $F$, while playing $S$ can expect $3(1-\lambda)$.
2. Type 2 **W** has expected payoff by playing $F$ $3(1-\lamba)$, while playing $S$ can expect $\lambda$.

## Example 2: Mixed Strategies

1. **W** type 1 would play a mixed strategies equilibrium only if $\lambda=3(1-\lambda)$, or $\lambda=\frac{3}{4}$.
2. **W** type 2 would play a mixed strategies equilibrium only if $3(1-\lambda)=\lambda$, or $\lambda=\frac{3}{4}$.

## Example 2: Mixed Strategies



# Mechanism Design, the Revelation Principle, and Sales to an Unknown Buyer

# Auctions

# Signaling Games and The Lemons Problem



# Bibliography

## Bibliography

These slides took definitions and examples from the book:

:::{.nonincremental}

* Strategies and Games: Theory and Applications from Prajit Dutta. 2001 MIT Press. Chapters 19 to 24.
* The Moral Hazard example was taken from *Wealth and the Principal-Agent Matching*, by P. Fagandini (2022), Managerial and Decision Economics.

:::